---
title: "Final Exam on Scorpion Behavior"
author: "Cora Peeler"
date: "05/26/2016"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(mosaic)
require(UsingR)
require(tigerstats)

```

# Introduction

<In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?>

#### what was asked for
Investigate whether the odors from a predator might influence the defensive behavior of Hadrurus arizonensis, and if the scorpion size plays a role in the type of response.

we see this as two questions

##### does presence of odor change the expected outcome on our measurements?

<for continuous measurements we could ask for center differences (means or medians)

for catagorical explanitory variables (such as count of type of response)
a chi-squared test would be chosen.>

We were given two comma separated values files containing these measurements:

* **Scorpion** - a tag indicating which scorpion was measured
* **Behavior** - a factor variable (no sting, dry sting, wet sting, jab)
* **Volume** - a numerical measure of venom  amount (wet sting)
* **Probes** - a numerical count of the number of "hits" applied to get a reaction
* **Reaction_Time** - a numerical measure of time it took to get a reaction
* **size** - a factor (large, small) of scorpion size
* **role** - a factor (control, threat) of absence or presence of an apparant threat to the scoprion (rat, urine)

There were 30 different scorpions tested. Each scorpion was tested twice, once in **control** (no rat urine present) and once in **threat** (rat urine). The order of testing was randomly assigned.

One of the csv files refers to the individual tests run. It has 60 observations. The others conveinently organized as paired data with observations of each scorpion in both threat and control conditions for conveinence in paired tests.

##### does size of the scorpion change the expected outcome on our measurements?

<for continuous measurements we could ask for center differences (means or medians)

for catagorical explanitory variables (such as count of type of response)
a chi-squared test would be chosen.>


# Methods for presence vs absence of odor

<Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.>


#Results

<Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.>

## Descriptive Results
In order to understand the comparison between the scorpions when odor is a factor to explain when they would attack or become more aggrevated. We must create a plot in order to compare the reaction time for the scorpions when they feel threatened by a predator and when they do not feel threatened by a predator. We must also produce a t.test in order to show the comparison between the control group as well as the experimental group so that we may see if they have a significant comparison.
### Graphical Descriptive Results

```{r}
url2="scorpion_reaction_rolls.csv"
scorp_data2=read.csv(url2,header=T)
contTable <- xtabs(~ Behavior + roll, data = scorp_data2)
barchart(contTable,
	auto.key = list(
		space = "top",
		title = "roll",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	main = "Behavior",
	xlab = "Roll",
	ylab = "Frequency")
```

The bar graph clearly shows that the scorpions are much more likely to have a wet sting when a threat is present, than in the control group when the threat is not present.

This plot shows that when we factor the reaction time for both the control and experimental group the comparison has a positive correlation which would indicate that the reaction time for the experiemental group has a direct correlation to the reaction time for the control group.

```{r}
bwplot(Reaction_Time ~ roll, data = scorp_data2, main = "Reaction Time by roll")
```

Looking at the boxplot, our eyes tell us that it looks like Reaction time under threat seems to be quicker and more trigger happen then opposed to the reaction time under control which looks slower.

<Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question>

### Numerical Descriptive Results

```{r}
attach(scorp_data2)
rol=table(roll,Behavior)
rowPerc(rol)
```

During the rowPerc test it clearly shows that the scorpions are more likely to have a wet sting when a threat is present as opposed to when there is no threat present. About 70% of the scorpions will have a wet sting when the threat is present; while those who did not have a threat present only about 7% had a wet sting.
<Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question>

## Inferential Results

```{r}
chisqtestGC(rol,simulate.p.value = TRUE ,graph=TRUE)
```

The chisqtest shows that we would reject the null hypothesis based on the fact that the wet sting was used more frequently with the threat prescence in comparison to when there is no threat present. The pvalue is significantly low which indicates that there is a difference between the two variables compared.

```{r}
t.test(Reaction_Time ~ roll, data = scorp_data2)
```

Looking at the group control we see that the average reaction time is 2.31 compared to the reaction time average for group threat of 1.15. Which is about 2 in 100. 

<State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question>

# Methods for different sizes (large vs small)

<Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.>

#Results

<Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.>

## Descriptive Results

### Graphical Descriptive Results

<Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question>



```{r}
url2="scorpion_reaction_size.csv"
scorp_data2=read.csv(url2,header=T)
contTable <- xtabs(~ Behavior + size, data = scorp_data2)
barchart(contTable,
	auto.key = list(
		space = "top",
		title = "Size",
		cex.title = 1,
		columns = 1),
	horizontal = FALSE,
	main = "Behavior",
	xlab = "Size",
	ylab = "Frequency")
```




```{r}
bwplot(Reaction_Time ~ size, data = scorp_data2, main = "Reaction Time by size")
```



### Numerical Descriptive Results


```{r}
attach(scorp_data2)
rol=table(size,Behavior)
rowPerc(rol)
```





<Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question>

## Inferential Results




```{r}
t.test(Reaction_Time ~ size, data = scorp_data2)
```



<State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question>


# Discussion and Conclusion for both set of research questions

<Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.>
